---
title: "4 x 2 Mixed ANOVA"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: "Guy A. Prochilo"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

# Required packages

```{r warning = FALSE, message = FALSE, echo = TRUE}
library(tidyverse) # Tidyverse package family
library(afex)
library(psych)
library(car)
library(emmeans)
library(ggpubr)
library(ggrepel)
library(cowplot)
```

# Analysis

## Import simulated dataset 

```{r}
# Import data and reshape for analysis
(read_csv("mixed_anova_dat.csv", col_types = cols()) %>% 
  separate(group, into = c("group", "time")) %>% 
  mutate(group = factor(group), 
         time = factor(time)) %>% 
  split(.$time) %>% 
  map(~mutate(., id = factor(1:length(.$time)))) %>% 
  bind_rows() %>% 
  select(id, group, time, dv) -> dat)
```

### Set `emmeans()` to use multivariate model for follow-up tests

```{r}
afex_options(emmeans_model = "multivariate")
```

### Analysis

#### Run the mixed ANOVA

```{r}
(dat %>% 
afex::aov_car(dv ~ group*time + Error(id/time), data = .,
              anova_table = list(es = c("pes", "ges"))) -> model)
```

#### Compute effect size for main effect and interaction

```{r}
(sjstats::anova_stats(model$aov) -> efx)
```

#### Compute descriptive statistics

```{r}
(model$data$long %>% 
  group_by(group, time) %>%
  summarise_at("dv", list(
    count = ~n(),
    mean = ~mean(.), 
    sd = ~sd(.),
    min = ~min(.),
    max = ~max(.),
    skew = ~psych::skew(.), 
    kurtosis = ~psych::kurtosi(.),
    shapiro_p = ~shapiro.test(.)$p.value)) -> descrip)
```

#### Plot descriptive data

```{r}
# Histograms and density plots of cell means
model$data$long %>% 
  ggplot(aes(dv)) +
  geom_histogram(aes(fill = group, y = ..density..), bins = 5, color = "black") +
  geom_density() +
  facet_wrap(group~time, scales = "free_x", nrow = 2)

# Data points, confidence intervals, and density plots of cell means
model$data$long %>% 
  ggplot(aes(y = dv, x = group)) +
  geom_point(aes(fill = group), pch = 21, position = position_jitterdodge()) +
  geom_violin(alpha = 0) +
  stat_summary(geom = "point", fun.y = mean, color = "black") +
  stat_summary(geom = "errorbar", fun.data = mean_cl_normal, width = 0.3) +
  facet_wrap(~time, scales = "free_x") +
  theme(legend.position = "none")
```

#### Compute Marginal Means of all cells in design

```{r}
(emm = emmeans(model, spec = ~group*time))
```

#### Compute Marginal Means for Main Effects

```{r}
# Main effect of time
(emm_time = emmeans(model, specs = ~time))

# Main effect of group
(emm_group = emmeans(model, specs = ~group))
```

#### Plot marginal means

```{r}
emm %>% 
  summary() %>% 
  {
    { . -> tmp} %>% 
  as.data.frame() %>% 
  {ggplot(., aes(x = factor(time),
                 y = emmean, 
                 group = factor(group),
                 shape = group)) +
      geom_point(position = position_dodge(0.2), size = 2) +
      geom_line(aes(linetype = group), position = position_dodge(0.2)) +
      geom_errorbar(inherit.aes = TRUE,
                    ymin = .$lower.CL, 
                    ymax = .$upper.CL,
                    width = 0.1, 
                    position = position_dodge(-0.2)) +
      theme_light(base_size = 15) + 
      theme(legend.position = "bottom", plot.caption = element_text(hjust = 0)) +
      labs(x = "Time", y = "DV", shape = "Group", linetype = "Group",
           caption = paste("-", attributes(tmp)$mesg, collapse = "\n")) +
      scale_x_discrete(labels = c("T0", "T1")) +
      scale_linetype_manual(values = c(1, 2, 3, 4), 
                            labels = c("Control", "Exercise", "MAP", "Meditation")) +
      scale_shape_manual(values = c(16, 17, 18, 19), 
                         labels = c("Control", "Exercise", "MAP", "Meditation")) +
      scale_y_continuous(limits = c(min(.$lower.CL), max(.$upper.CL)),
                   breaks = scales::pretty_breaks(n = 10))}
  }
```

#### Follow up interaction using comparisons of time within each group

```{r}
(emm %>% 
  contrast(., method = "revpairwise", by = "group") %>% 
  summary(., 
          infer = c(T, T), 
          level = 0.95, 
          null = 0, 
          by = NULL, adjust = "bonf") -> int)

int %>% 
  {plot(., horizontal = F, CIs = F, colors = "black") +
      theme_light(base_size = 15) + 
      geom_hline(yintercept = 0, linetype = "dashed") +
      geom_errorbar(inherit.aes = TRUE,
                    ymin = .$lower.CL,
                    ymax = .$upper.CL,
                    width = 0.05) +
      ylim(min(.$lower.CL, 0), max(.$upper.CL, 0)) +
      labs(caption = paste("-", attributes(.)$mesg, collapse = "\n")) +
      xlab(label = "Time") + 
      ylab(label = "DV") +
      labs(y = "DV", x = "Contrast") +
      theme(legend.position = "bottom", plot.caption = element_text(hjust = 0))}
```

#### Determine if there is a difference between the t1 -- t0 gain scores for MAP training compared to all other training modalities

```{r}
# Define full contrast matrix for ease of use
cont1 = c(1, 0, 0, 0, 0, 0, 0, 0)
cont2 = c(0, 1, 0, 0, 0, 0, 0, 0)
ext1 = c(0, 0, 1, 0, 0, 0, 0, 0)
ext2 = c(0, 0, 0, 1, 0, 0, 0, 0)
mapt1 = c(0, 0, 0, 0, 1, 0, 0, 0)
mapt2 = c(0, 0, 0, 0, 0, 1, 0, 0)
medt1 = c(0, 0, 0, 0, 0, 0, 1, 0)
medt2 = c(0, 0, 0, 0, 0, 0, 0, 1)

# Compute contrasts
(emmeans(model, spec = ~time*group) %>% 
  contrast(., method = list("map(t2-t1) - con(t2-t1)" = ((mapt2 - mapt1) - (cont2 - cont1)), 
                            "map(t2-t1) - ex(t2-t1)" = ((mapt2 - mapt1) - (ext2 - ext1)), 
                            "map(t2-t1) - med(t2-t1)"= ((mapt2 - mapt1) - (medt2 - medt1)))) %>% 
  summary(., infer = c(T, T), level = 0.95, null = 0, adjust = "bonf") -> cons)

# Plot contrasts
cons %>% 
  {plot(., horizontal = F, CIs = F, colors = "black") +
      theme_light(base_size = 15) + 
      geom_hline(yintercept = 0, linetype = "dashed") +
      geom_errorbar(inherit.aes = TRUE,
                    ymin = .$lower.CL,
                    ymax = .$upper.CL,
                    width = 0.05) +
      ylim(min(.$lower.CL, 0), max(.$upper.CL, 0)) +
      labs(caption = paste("-", attributes(.)$mesg, collapse = "\n")) +
      xlab(label = "Time") + 
      ylab(label = "DV") +
      labs(y = "DV", x = "Contrast") +
      theme(legend.position = "bottom", plot.caption = element_text(hjust = 0))}
```

